---
title: "TroutTime"
author: "Dylan Rees Jones"
date: "2026-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# install.packages(c("rnaturalearth", "rnaturalearthdata", "sf"))
# install.packages("ggpattern")
library(readr)
library(VennDiagram)
library(tidyr)
records <- read_csv("WilderlabPublicData-2026-08-17 1403/records.csv")
samples <- read_csv("WilderlabPublicData-2026-08-17 1403/samples.csv")
```
##**First lookin around**
```{r}
nrow(records) # <- Checking the number of samples
nrow(samples) # <- ""

# Setting up different trout data
Trout <- records[records$CommonName == "Trout", ]
BrownTrout <- records[records$CommonName == "Brown Trout", ]
RainTrout <- records[records$CommonName == "Rainbow Trout", ]

# Includes all the other kind of trout thingos
allTrout <- records[grepl("trout", records$CommonName, ignore.case = TRUE), ]

# Kind of weird, not organised amazingly well
table(allTrout$CommonName)

## Gonna stick with just brown trout and rainbow trout
duoTrout <- rbind(BrownTrout, RainTrout)
```


```{r}
## Rearanging the dataframe

# How many IDs there are
length(unique(records$UID))

# Makes the dataframes
commonNamesList <- as.list(unique(records$CommonName))
uidsList <- as.list(unique(records$UID))

orgDF <- data.frame(matrix(NA, nrow = length(uidsList), ncol = length(commonNamesList)))
rownames(orgDF) <- uidsList
colnames(orgDF) <- commonNamesList

# Checks how many times each fish shows up in the data
# Gonna keep em all
fishPlot <- table(records$CommonName)
barplot(fishPlot, las = 2)

# Reorders the dataframe
records <- records[!is.na(records$CommonName), ]
for (i in 1:nrow(records)) {
  id <- records$UID[i]
  name <- records$CommonName[i]
  count <- records$Count[i]
  orgDF[as.character(id), as.character(name)] <- count
}
```

# Amount of trout
```{r}
yesRainTroutUIDs <- orgDF[!is.na(orgDF$"Rainbow trout"), ]
noRainTroutUIDs <- orgDF[is.na(orgDF$"Rainbow trout"), ]
yesBrownTroutUIDs <- orgDF[!is.na(orgDF$"Brown trout"), ]
noBrownTroutUIDs <- orgDF[is.na(orgDF$"Brown trout"), ]
yesGalxiidsUIDs <- orgDF[is.na(orgDF$"Galaxiid"), ]
noGalxiidsUIDs <- orgDF[!is.na(orgDF$"Galaxiid"), ]

troutVar1 <- intersect(rownames(noRainTroutUIDs), rownames(noBrownTroutUIDs))
troutfullAll_UIDs <- noRainTroutUIDs[troutVar1, ]

troutVar2 <- intersect(rownames(yesRainTroutUIDs), rownames(yesBrownTroutUIDs))
troutlessAll_UIDs <- yesRainTroutUIDs[troutVar2, ]
```


# Co-occurence of trout
```{r}
venn.diagram(x = list(rownames(noRainTroutUIDs), rownames(noBrownTroutUIDs)),
             category.names = c("NoRainTrout", "NoBrownTrout"),
             filename = "vennNo.png",
             imagetype = "png")

venn.diagram(x = list(rownames(yesRainTroutUIDs), rownames(yesBrownTroutUIDs)),
             category.names = c("YesRainTrout", "YesBrownTrout"),
             filename = "vennYes.png",
             imagetype = "png")

## Chi squared test
# Grabs just the rainbow and brown trout
rainBrownOrgDF <- orgDF[, c("Rainbow trout", "Brown trout")]
rainBrownOrgDF[is.na(rainBrownOrgDF)] <- 0

# Does the test
tab <- table(rainBrownOrgDF$"Rainbow trout" > 0, rainBrownOrgDF$"Brown trout" > 0)
model <- chisq.test(tab)
model

# actual counts
model$observed
# counts expected if independent
model$expected
```

```{r}
## Trout abundance (Contains only samples where at least one of the trout were present)
library(tidyr)
library(ggplot2)

troutDF <- rainBrownOrgDF
colnames(troutDF) <- c("RainbowTrout", "BrownTrout")

# Keep only samples with at least one trout present
troutDF <- troutDF[troutDF$RainbowTrout > 0 | troutDF$BrownTrout > 0, ]

# Determine which species is higher per sample
troutDF$Higher <- ifelse(troutDF$RainbowTrout > troutDF$BrownTrout, "RainbowTrout", "BrownTrout")

# Reorder: BrownTrout group first, then RainbowTrout group
troutDF <- troutDF[order(factor(troutDF$Higher, levels = c("BrownTrout", "RainbowTrout"))), ]

# Use rownames as UID, preserving order via factor levels
troutDF$UID <- factor(rownames(troutDF), levels = rownames(troutDF))

troutLong <- pivot_longer(troutDF, cols = c("RainbowTrout", "BrownTrout"),
                           names_to = "Species", values_to = "Count")
troutLong <- troutLong[troutLong$Count > 0, ]

ggplot() +
  geom_vline(data = troutDF, aes(xintercept = as.numeric(UID), color = Higher),
             linewidth = 3, alpha = 0.15, show.legend = FALSE) +
  geom_point(data = troutLong, aes(x = as.numeric(UID), y = Count, fill = Species),
             shape = 21, color = "black", size = 2.5, stroke = 0.5) +
  scale_color_manual(values = c("RainbowTrout" = "#90c990", "BrownTrout" = "#c8a27a")) +
  scale_fill_manual(values = c("RainbowTrout" = "#1a8c1a", "BrownTrout" = "#8b5a2b")) +
  labs(title = "Trout Abundance per Sample", x = "Sample (UID)", y = "Count") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```
Okay, thats cool, but theres the crazy outliers, gonna leave this plot how it is but try make a new one without them.
Then gonna look at which areas these were and try figure out why tf they are like that

```{r}
troutDF2 <- rainBrownOrgDF
colnames(troutDF2) <- c("RainbowTrout", "BrownTrout")

# Keep only samples with at least one trout present
troutDF2 <- troutDF2[troutDF2$RainbowTrout > 0 | troutDF2$BrownTrout > 0, ]

# Determine which species is higher per sample
troutDF2$Higher <- ifelse(troutDF2$RainbowTrout > troutDF2$BrownTrout, "RainbowTrout", "BrownTrout")

# Reorder: BrownTrout group first, then RainbowTrout group
troutDF2 <- troutDF2[order(factor(troutDF2$Higher, levels = c("BrownTrout", "RainbowTrout"))), ]

# Use rownames as UID, preserving order via factor levels
troutDF2$UID <- factor(rownames(troutDF2), levels = rownames(troutDF2))

troutLong2 <- pivot_longer(troutDF2, cols = c("RainbowTrout", "BrownTrout"),
                            names_to = "Species", values_to = "Count")
troutLong2 <- troutLong2[troutLong2$Count > 0, ]

# Remove points with count higher than 300000
troutLong2 <- troutLong2[troutLong2$Count <= 300000, ]

ggplot() +
  geom_vline(data = troutDF2, aes(xintercept = as.numeric(UID), color = Higher),
             linewidth = 3, alpha = 0.15, show.legend = FALSE) +
  geom_point(data = troutLong2, aes(x = as.numeric(UID), y = Count, fill = Species),
             shape = 21, color = "black", size = 2.5, stroke = 0.5) +
  scale_color_manual(values = c("RainbowTrout" = "#90c990", "BrownTrout" = "#c8a27a")) +
  scale_fill_manual(values = c("RainbowTrout" = "#1a8c1a", "BrownTrout" = "#8b5a2b")) +
  labs(title = "Trout Abundance per Sample", x = "Sample (UID)", y = "Count") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

```{r}
# Rebuild full long data (before outlier removal) to identify what got removed
troutLongFull2 <- pivot_longer(troutDF2, cols = c("RainbowTrout", "BrownTrout"),
                                names_to = "Species", values_to = "Count")
troutLongFull2 <- troutLongFull2[troutLongFull2$Count > 0, ]

# Outliers = points removed by the >300000 filter
outliers2 <- troutLongFull2[troutLongFull2$Count > 300000, ]
outliers2

# UIDs of outlier samples
outlierUIDs2 <- unique(as.character(outliers2$UID))
outlierUIDs2

# Pull matching sample info from samples.csv
outlierSamples2 <- samples[samples$UID %in% outlierUIDs2, ]
outlierSamples2
```
Plots the outliers on NZ map
```{r}
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)

# Get species info (Higher) attached to each outlier UID
outlierInfo2 <- merge(outliers2, samples, by = "UID")

# Get NZ map
nz <- ne_countries(scale = "medium", country = "New Zealand", returnclass = "sf")

ggplot() +
  geom_sf(data = nz, fill = "grey95", color = "black") +
  geom_point(data = outlierInfo2, aes(x = Longitude, y = Latitude, fill = Species),
             shape = 21, color = "black", size = 3, stroke = 0.5) +
  scale_fill_manual(values = c("RainbowTrout" = "#1a8c1a", "BrownTrout" = "#8b5a2b")) +
  coord_sf(xlim = c(166, 179), ylim = c(-47.5, -34)) +
  labs(title = "Locations of Outlier Trout Samples", x = "Longitude", y = "Latitude") +
  theme_minimal()
```



Plotting only where there are rainbow and brown trout
```{r}
troutDF3 <- rainBrownOrgDF
colnames(troutDF3) <- c("RainbowTrout", "BrownTrout")

# Keep only samples with BOTH rainbow and brown trout present
troutDF3 <- troutDF3[troutDF3$RainbowTrout > 0 & troutDF3$BrownTrout > 0, ]

# Determine which species is higher per sample
troutDF3$Higher <- ifelse(troutDF3$RainbowTrout > troutDF3$BrownTrout, "RainbowTrout", "BrownTrout")

# Reorder: BrownTrout group first, then RainbowTrout group
troutDF3 <- troutDF3[order(factor(troutDF3$Higher, levels = c("BrownTrout", "RainbowTrout"))), ]

# Use rownames as UID, preserving order via factor levels
troutDF3$UID <- factor(rownames(troutDF3), levels = rownames(troutDF3))

troutLong3 <- pivot_longer(troutDF3, cols = c("RainbowTrout", "BrownTrout"),
                            names_to = "Species", values_to = "Count")

# Remove points with count higher than 300000
troutLong3 <- troutLong3[troutLong3$Count <= 300000, ]

ggplot() +
  geom_vline(data = troutDF3, aes(xintercept = as.numeric(UID), color = Higher),
             linewidth = 3, alpha = 0.15, show.legend = FALSE) +
  geom_point(data = troutLong3, aes(x = as.numeric(UID), y = Count, fill = Species),
             shape = 21, color = "black", size = 2.5, stroke = 0.5) +
  scale_color_manual(values = c("RainbowTrout" = "#90c990", "BrownTrout" = "#c8a27a")) +
  scale_fill_manual(values = c("RainbowTrout" = "#1a8c1a", "BrownTrout" = "#8b5a2b")) +
  labs(title = "Trout Abundance per Sample (Both Species Present)", x = "Sample (UID)", y = "Count") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

```{r}
# Pie chart
library(ggplot2)

pieDF <- rainBrownOrgDF
colnames(pieDF) <- c("RainbowTrout", "BrownTrout")

pieDF$Category <- ifelse(pieDF$RainbowTrout > 0 & pieDF$BrownTrout > 0, "Both",
                   ifelse(pieDF$RainbowTrout > 0, "RainbowOnly",
                   ifelse(pieDF$BrownTrout > 0, "BrownOnly", NA)))

pieDF <- pieDF[!is.na(pieDF$Category), ]

pieCounts <- as.data.frame(table(pieDF$Category))
colnames(pieCounts) <- c("Category", "Count")
pieCounts$Percent <- round(100 * pieCounts$Count / sum(pieCounts$Count), 1)

ggplot(pieCounts, aes(x = "", y = Count, fill = Category)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(Percent, "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("Both" = "#a3a3c2", "RainbowOnly" = "#1a8c1a", "BrownOnly" = "#8b5a2b")) +
  theme_void() +
  labs(title = "Trout Presence Breakdown", fill = "Category")
```

```{r}
# Categorize each sample based on trout presence
mapDF <- rainBrownOrgDF
colnames(mapDF) <- c("RainbowTrout", "BrownTrout")
mapDF$UID <- rownames(mapDF)
mapDF$Category <- ifelse(mapDF$RainbowTrout > 0 & mapDF$BrownTrout == 0, "RainbowOnly",
                   ifelse(mapDF$BrownTrout > 0 & mapDF$RainbowTrout == 0, "BrownOnly",
                   ifelse(mapDF$RainbowTrout == 0 & mapDF$BrownTrout == 0, "Neither", "Both")))

# Merge with sample coordinates
mapInfo <- merge(mapDF, samples, by = "UID")
mapInfo <- mapInfo[mapInfo$Category != "Neither", ]

ggplot() +
  geom_sf(data = nz, fill = "grey95", color = "black") +
  geom_point(data = mapInfo, aes(x = Longitude, y = Latitude, fill = Category),
             shape = 21, color = "black", size = 1.5, stroke = 0.3, alpha = 0.5) +
  scale_fill_manual(values = c("RainbowOnly" = "#1a8c1a",
                                "BrownOnly" = "#8b5a2b",
                                "Both" = "#a3a3c2")) +
  guides(fill = guide_legend(override.aes = list(size = 4, alpha = 1))) +
  coord_sf(xlim = c(166, 179), ylim = c(-47.5, -34)) +
  labs(title = "Trout Presence by Sample Site", x = "Longitude", y = "Latitude") +
  theme_minimal()
```

```{r}
## Merge with samples first to get Latitude
mapDF <- merge(rainBrownOrgDF, samples[, c("UID", "Latitude", "Longitude")],
               by.x = "row.names", by.y = "UID")
colnames(mapDF)[1] <- "UID"
colnames(mapDF)[2:3] <- c("RainbowTrout", "BrownTrout")

# Remove corrupted coordinates
mapDF <- mapDF[!(mapDF$UID %in% c("744321","744322","744323","744324","744325","744326")), ]

mapDF$Category <- ifelse(mapDF$RainbowTrout > 0 & mapDF$BrownTrout == 0, "RainbowOnly",
                   ifelse(mapDF$BrownTrout > 0 & mapDF$RainbowTrout == 0, "BrownOnly",
                   ifelse(mapDF$RainbowTrout == 0 & mapDF$BrownTrout == 0, "Neither", "Both")))

# Final island classification rule
mapDF$Island <- ifelse(
  mapDF$Latitude > -40.2, "North Island",
  ifelse(mapDF$Longitude < 174.3, "South Island", "North Island")
)

# Map plot
ggplot() +
  geom_sf(data = nz, fill = "grey95", color = "black") +
  geom_point(data = mapDF, aes(x = Longitude, y = Latitude, color = Island),
             size = 1.5, alpha = 0.5) +
  scale_color_manual(values = c("North Island" = "red", "South Island" = "blue")) +
  coord_sf(xlim = c(166, 179), ylim = c(-47.5, -34)) +
  labs(title = "Samples by Island", x = "Longitude", y = "Latitude") +
  theme_minimal()

# Summarize counts per island
pieData <- as.data.frame(table(mapDF$Island, mapDF$Category))
colnames(pieData) <- c("Island", "Category", "Count")
pieData <- pieData[pieData$Count > 0, ]

pieData <- do.call(rbind, lapply(split(pieData, pieData$Island), function(d) {
  d$Percent <- round(100 * d$Count / sum(d$Count), 1)
  d
}))

pieDataNorth <- pieData[pieData$Island == "North Island", ]
pieDataSouth <- pieData[pieData$Island == "South Island", ]

# North Island pie
pieDataNorth$Label <- paste0(pieDataNorth$Category, ": ", pieDataNorth$Count, " (", pieDataNorth$Percent, "%)")
colorMap <- c("RainbowOnly" = "#1a8c1a", "BrownOnly" = "#8b5a2b", "Both" = "#a3a3c2", "Neither" = "grey60")
pieDataNorth$Color <- colorMap[pieDataNorth$Category]

ggplot(pieDataNorth, aes(x = "", y = Count, fill = Label)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = setNames(pieDataNorth$Color, pieDataNorth$Label)) +
  theme_void() +
  labs(title = "North Island: Trout Presence", fill = "Category")

# South Island pie
pieDataSouth$Label <- paste0(pieDataSouth$Category, ": ", pieDataSouth$Count, " (", pieDataSouth$Percent, "%)")
pieDataSouth$Color <- colorMap[pieDataSouth$Category]

ggplot(pieDataSouth, aes(x = "", y = Count, fill = Label)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = setNames(pieDataSouth$Color, pieDataSouth$Label)) +
  theme_void() +
  labs(title = "South Island: Trout Presence", fill = "Category")
```

##**Diversity**
```{r}
diversityDF$Category2 <- ifelse(diversityDF$RainbowTrout > 0 & diversityDF$BrownTrout == 0, "RainbowOnly",
                          ifelse(diversityDF$BrownTrout > 0 & diversityDF$RainbowTrout == 0, "BrownOnly",
                          ifelse(diversityDF$RainbowTrout == 0 & diversityDF$BrownTrout == 0, "Neither",
                          "Both")))

# Build a long version with an extra "Either" group added alongside the four categories
eitherDF <- diversityDF[diversityDF$RainbowTrout > 0 | diversityDF$BrownTrout > 0, ]
eitherDF$Category2 <- "Either"

diversityLong <- rbind(diversityDF, eitherDF)
diversityLong$Category2 <- factor(diversityLong$Category2,
                                   levels = c("RainbowOnly", "BrownOnly", "Both", "Neither", "Either"))

# Richness boxplot
ggplot(diversityLong, aes(x = Category2, y = Richness, fill = Category2)) +
  geom_boxplot() +
  scale_fill_manual(values = c("RainbowOnly" = "#1a8c1a", "BrownOnly" = "#8b5a2b",
                                "Both" = "#a3a3c2", "Neither" = "grey60", "Either" = "#d9a441")) +
  labs(title = "Species Richness by Trout Presence", y = "Species Richness", x = "Trout Category") +
  theme_minimal()

# Shannon diversity boxplot
ggplot(diversityLong, aes(x = Category2, y = Shannon, fill = Category2)) +
  geom_boxplot() +
  scale_fill_manual(values = c("RainbowOnly" = "#1a8c1a", "BrownOnly" = "#8b5a2b",
                                "Both" = "#a3a3c2", "Neither" = "grey60", "Either" = "#d9a441")) +
  labs(title = "Shannon Diversity by Trout Presence", y = "Shannon Index", x = "Trout Category") +
  theme_minimal()
```
WHEN TROUT ARRIVE DOES IT FUCK THE DIVERSITY

```{r}
analyze_trout_arrival <- function(fullDF, troutPresentFunc, label) {

  fullDF$TroutPresent <- troutPresentFunc(fullDF)

  siteCounts <- table(fullDF$SiteID)
  validSites <- names(siteCounts[siteCounts >= 2])
  siteDF <- fullDF[fullDF$SiteID %in% validSites, ]

  siteDF <- siteDF[order(siteDF$SiteID, siteDF$CollectionDate), ]

  results <- do.call(rbind, lapply(split(siteDF, siteDF$SiteID), function(d) {
    first <- d[1, ]
    last <- d[nrow(d), ]
    data.frame(
      SiteID = first$SiteID,
      Latitude = first$Latitude,
      Longitude = first$Longitude,
      TroutFirst = first$TroutPresent,
      TroutLast = last$TroutPresent,
      ShannonFirst = first$Shannon,
      ShannonLast = last$Shannon,
      DiversityChange = last$Shannon - first$Shannon,
      DateFirst = first$CollectionDate,
      DateLast = last$CollectionDate,
      YearsGap = as.numeric(difftime(last$CollectionDate, first$CollectionDate, units = "days")) / 365.25
    )
  }))

  arrivalSites <- results[results$TroutFirst == FALSE & results$TroutLast == TRUE, ]
  return(arrivalSites)
}

# Regenerate arrival datasets with date info
rainbowArrival <- analyze_trout_arrival(fullDF, function(d) d$RainbowTrout > 0, "Rainbow Trout")
brownArrival <- analyze_trout_arrival(fullDF, function(d) d$BrownTrout > 0, "Brown Trout")
bothArrival <- analyze_trout_arrival(fullDF, function(d) d$RainbowTrout > 0 & d$BrownTrout > 0, "Both Trout")
eitherArrival <- analyze_trout_arrival(fullDF, function(d) d$RainbowTrout > 0 | d$BrownTrout > 0, "Either Trout")

library(patchwork)

plot_arrival_single <- function(arrivalData, label, gap) {
  filtered <- arrivalData[arrivalData$YearsGap >= gap, ]

  maxAbs <- max(abs(filtered$DiversityChange), na.rm = TRUE)

  ggplot() +
    geom_sf(data = nz, fill = "grey95", color = "black") +
    geom_point(data = filtered, aes(x = Longitude, y = Latitude, fill = DiversityChange),
               shape = 21, color = "black", size = 2, alpha = 0.9, stroke = 0.3) +
    scale_fill_gradient2(low = "red", mid = "white", high = "green", midpoint = 0,
                          limits = c(-maxAbs/4, maxAbs/4), oob = scales::squish) +
    coord_sf(xlim = c(166, 179), ylim = c(-47.5, -34)) +
    labs(title = paste0(label, " (n=", nrow(filtered), ")"), fill = "Shannon\nChange") +
    theme_minimal() +
    theme(axis.title = element_blank(), axis.text = element_text(size = 6))
}

gaps <- c(0, 1, 2)

for (g in gaps) {
  p1 <- plot_arrival_single(rainbowArrival, "Rainbow Trout", g)
  p2 <- plot_arrival_single(brownArrival, "Brown Trout", g)
  p3 <- plot_arrival_single(bothArrival, "Both Trout", g)
  p4 <- plot_arrival_single(eitherArrival, "Either Trout", g)

  combined <- (p1 | p2) / (p3 | p4) +
    plot_annotation(title = paste0("Diversity Change on Trout Arrival (", g, "+ yr gap)"))

  print(combined)
}
```
WHEN TROUT INCREASES DOES IT FUCK THE DIVERSITY
```{r}
# Recompute using ALL sites with 2+ samples
siteCounts2 <- table(fullDF$SiteID)
validSites2 <- names(siteCounts2[siteCounts2 >= 2])
siteDF2 <- fullDF[fullDF$SiteID %in% validSites2, ]
siteDF2 <- siteDF2[order(siteDF2$SiteID, siteDF2$CollectionDate), ]

changeDF2 <- do.call(rbind, lapply(split(siteDF2, siteDF2$SiteID), function(d) {
  first <- d[1, ]
  last <- d[nrow(d), ]
  data.frame(
    SiteID = first$SiteID,
    Latitude = first$Latitude,
    Longitude = first$Longitude,
    TroutChange_Rainbow2 = last$RainbowTrout - first$RainbowTrout,
    TroutChange_Brown2 = last$BrownTrout - first$BrownTrout,
    TroutChange_Both2 = pmin(last$RainbowTrout, last$BrownTrout) - pmin(first$RainbowTrout, first$BrownTrout),
    TroutChange_Either2 = pmax(last$RainbowTrout, last$BrownTrout) - pmax(first$RainbowTrout, first$BrownTrout),
    DiversityChange2 = last$Shannon - first$Shannon,
    YearsGap2 = as.numeric(difftime(last$CollectionDate, first$CollectionDate, units = "days")) / 365.25
  )
}))

# Plotting function: filters to sites where trout increased, colors by diversity change
plot_change_single2 <- function(changeDF2, troutCol, label, gap) {
  plotData <- changeDF2[changeDF2$YearsGap2 >= gap, ]
  plotData <- plotData[plotData[[troutCol]] > 0, ]

  maxAbs <- max(abs(plotData$DiversityChange2), na.rm = TRUE)

  ggplot() +
    geom_sf(data = nz, fill = "grey95", color = "black") +
    geom_point(data = plotData, aes(x = Longitude, y = Latitude, fill = DiversityChange2),
               shape = 21, color = "black", size = 2, alpha = 0.9, stroke = 0.3) +
    scale_fill_gradient2(low = "red", mid = "white", high = "green", midpoint = 0,
                          limits = c(-maxAbs/4, maxAbs/4), oob = scales::squish) +
    coord_sf(xlim = c(166, 179), ylim = c(-47.5, -34)) +
    labs(title = paste0(label, " (n=", nrow(plotData), ")"), fill = "Shannon\nChange") +
    theme_minimal() +
    theme(axis.title = element_blank(), axis.text = element_text(size = 6))
}

# Generate combined plots for each year-gap threshold
gaps2 <- c(0, 1, 2)

for (g in gaps2) {
  p1 <- plot_change_single2(changeDF2, "TroutChange_Rainbow2", "Rainbow Trout Increase", g)
  p2 <- plot_change_single2(changeDF2, "TroutChange_Brown2", "Brown Trout Increase", g)
  p3 <- plot_change_single2(changeDF2, "TroutChange_Both2", "Both Trout Increase", g)
  p4 <- plot_change_single2(changeDF2, "TroutChange_Either2", "Either Trout Increase", g)

  combined <- (p1 | p2) / (p3 | p4) +
    plot_annotation(title = paste0("Diversity Change on Trout Increase (", g, "+ yr gap)"))

  print(combined)
}
```


```{r}
library(tidyr)
# Trout count change density plot, filtered to sites where that trout type increased
troutChangeLong2_filtered <- troutChangeLong2[troutChangeLong2$TroutChangeValue > 0, ]

ggplot(troutChangeLong2_filtered, aes(x = TroutChangeValue, fill = TroutType)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Trout Count Increase", x = "Increase in Trout Count", y = "Density", fill = "Trout Type") +
  theme_minimal()

# Diversity change density plot, one curve per trout type, using sites where that type increased
diversityLong2_filtered <- pivot_longer(
  changeDF2,
  cols = c("TroutChange_Rainbow2", "TroutChange_Brown2", "TroutChange_Both2", "TroutChange_Either2"),
  names_to = "TroutType",
  values_to = "TroutChangeValue"
)

diversityLong2_filtered$TroutType <- factor(diversityLong2_filtered$TroutType,
  levels = c("TroutChange_Rainbow2", "TroutChange_Brown2", "TroutChange_Both2", "TroutChange_Either2"),
  labels = c("Rainbow", "Brown", "Both", "Either"))

diversityLong2_filtered <- diversityLong2_filtered[diversityLong2_filtered$TroutChangeValue > 0, ]

ggplot(diversityLong2_filtered, aes(x = DiversityChange2, fill = TroutType)) +
  geom_density(alpha = 0.4) +
  labs(title = "Diversity Change at Sites with Trout Increase", x = "Change in Shannon Diversity", y = "Density", fill = "Trout Type") +
  theme_minimal()
```


```{r}
siteCounts <- as.data.frame(table(fullDF$SiteID))
colnames(siteCounts) <- c("SiteID", "NumSamples")

# Number of sites with 2+ samples (repeat visits)
sum(siteCounts$NumSamples >= 2)

# Full breakdown
table(siteCounts$NumSamples)
```

